% Section 3.1.1 - Missing data paragraph was trimmed to only report the observation 
% of missingness (which fields, how much), with a forward reference to Section 3.2.1 
% for the treatment. The philosophical justification for listwise deletion was moved out.

% Section 3.1.2 - Renamed the subsubsection from "Pre-processing and Normalisation" 
% to "Upstream Processing and Normalisation" to make clear this only describes what 
% TCGA/GDC did, not what you did. The matrix transposition and data leakage paragraphs 
% were removed entirely, replaced by a single forward-reference sentence pointing to Section 3.2.

% Section 3.2.1 - Handling Residual Missing Values now contains the philosophical 
% justification for listwise deletion (moved from 3.1.1), which reads naturally 
% here because the numbers and decisions are already in context. I also added a 
% placeholder citation \cite{sterne2009multiple} for the missing data handling 
% rationale - you should replace this with whatever missing data reference you're 
% using, or remove it if you prefer not to cite one.

% Section 3.2.2 is unchanged in content but is now the only place where data 
% leakage is discussed, eliminating the duplication.

\typeout{NT FILE chapter3.tex}
% ============================================================
%  Chapter 3 - Methodology and Pilot Study
% ============================================================

\chapter{Methodology and Pilot Study}

Small introduction to the Chapter.

% ============================================================
\section{Data Acquisition and Characterisation}
% ============================================================

\subsection{The Primary Cohort (\textit{N}\,=\,272): Source and Clinical Features}
\label{subsec:primary-cohort}

The dataset used throughout this work was obtained from The Cancer Genome Atlas
Breast Invasive Carcinoma (TCGA-BRCA) study, accessed via the cBioPortal for
Cancer Genomics~\cite{https://www.cbioportal.org/study/summary?id=brca_tcga_pub}
under the identifier \texttt{brca\_tcga\_pub}. TCGA-BRCA is one of the most
comprehensive multi-omics repositories available for breast cancer research,
providing harmonised clinical annotations, genomic profiles, and molecular
expression data for a large cohort of patients profiled with standardised
protocols.

The full clinical release contains records for 463 patients. After restricting
the cohort to individuals for whom matched miRNA expression profiles were
available (see Section~\ref{subsec:mirna-profiles}), a final working cohort of
\textbf{272 samples} was obtained. All samples correspond to primary solid tumour
tissue (\texttt{SAMPLE\_TYPE = Primary}) from female patients, consistent with
the epidemiology of the disease. The clinical data matrix comprises 37 variables
per patient, covering receptor status, tumour staging, survival outcomes, and
several integrative molecular annotations.

\paragraph{Molecular subtype labels.}
The primary classification target in this thesis is the \textbf{PAM50 intrinsic
subtype}~\cite{parker2009supervised}, a well-established gene-expression-based
taxonomy that partitions breast cancer into five biologically distinct subtypes.
Table~\ref{tab:bc_subtypes_summary} summarises the key biological characteristics
of the four main subtypes; Normal-like is the fifth but is omitted from that
summary due to its low prevalence in the dataset. The subtype distribution
within the working cohort is presented in Table~\ref{tab:pam50-distribution}.
Luminal~A is the dominant class (43\%), whereas Normal-like is severely
underrepresented ($n$\,=\,5; 1.8\%). This pronounced class imbalance is a
central challenge addressed in the modelling strategy described in Chapter~4.

\begin{table}[h!]
    \centering
    \caption{PAM50 subtype distribution in the working cohort ($N$\,=\,272).}
    \label{tab:pam50-distribution}
    \begin{tabular}{lrr}
        \hline
        \textbf{PAM50 Subtype} & \textbf{Count} & \textbf{Proportion (\%)} \\
        \hline
        Luminal A      & 117 & 43.0 \\
        Luminal B      &  75 & 27.6 \\
        Basal-like     &  45 & 16.5 \\
        HER2-enriched  &  30 & 11.0 \\
        Normal-like    &   5 &  1.8 \\
        \hline
        \textbf{Total} & \textbf{272} & \textbf{100.0} \\
        \hline
    \end{tabular}
\end{table}

\paragraph{Demographic and clinical characteristics.}
Patient ages range from 26 to 90 years, with a mean of 57.9\,$\pm$\,13.3 years
(median\,=\,58). The age distribution is approximately symmetric around the
sixth decade of life, consistent with the known epidemiology of post-menopausal
breast cancer~\cite{Postmenopausal_bca_Rosenberg2006Risk}.

Receptor status was assessed immunohistochemically for the three canonical
markers. Oestrogen receptor (ER) positivity was recorded in 76.5\% of patients,
progesterone receptor (PR) positivity in 65.9\%, and HER2 amplification in
14.7\%. These proportions reflect the expected distribution in an unselected
breast cancer population and align with the PAM50 subtype assignments: the
Basal-like subclass is largely composed of triple-negative tumours
(ER$^{-}$/PR$^{-}$/HER2$^{-}$), a well-documented biological correspondence
that provides a useful sanity check for classifier outputs.

Tumour stage at diagnosis was available for 271 patients: stage T2 was the most
frequent (57.9\%), followed by T1 (25.9\%), T3 (11.7\%), and T4 (3.7\%). From
an overall survival perspective, 87.5\% of patients ($n$\,=\,238) were alive at
the time of last follow-up, with 58 recorded deaths.

\paragraph{Missing data.}
Several integrative annotation fields - specifically the
\texttt{INTEGRATED\_CLUSTERS} variables ($\approx$25\% missing) and the
\texttt{RPPA\_CLUSTER} field ($\approx$24\% missing) - contain substantial
proportions of missing values, reflecting the multi-platform nature of the
TCGA-BRCA data collection, where not every assay was performed on every sample.
Beyond these high-missingness columns, a small number of samples present
isolated missing values in fields such as \texttt{HER2\_STATUS},
\texttt{METASTASIS\_CODED}, and \texttt{TUMOR\_T1\_CODED}. The precise
treatment of all missing data is described in Section~\ref{subsec:missing-outliers}.
The PAM50 label, however, is complete across all 272 samples and was therefore
adopted as the sole ground-truth annotation for all supervised and unsupervised
experiments in this thesis.

% --------------------------------------------------------------
\subsection{miRNA Expression Profiles: Data Distribution and Normalisation}
\label{subsec:mirna-profiles}

\subsubsection{Dataset Dimensions and Data Type}

The miRNA expression matrix was obtained from the same TCGA-BRCA release and
encompasses \textbf{888 miRNA features} measured across \textbf{272 patient
samples}. This is a canonical example of a \textit{high-dimensional,
low-sample-size} (HDLSS) setting, where the number of features ($p = 888$)
greatly exceeds the number of observations ($n = 272$), posing well-known
challenges for statistical learning that motivate the feature selection
strategies explored in Chapter~4.

\subsubsection{Upstream Processing and Normalisation}

The expression values delivered by the TCGA data portal had already undergone
upstream processing by the GDC harmonisation pipeline: raw sequencing reads were
aligned to the reference genome, per-miRNA read counts were quantified, and
values were normalised and log$_2$-transformed. The presence of negative values
in the matrix (6.3\% of all entries, minimum\,=\,$-2.72$) is consistent with a
log$_2$-transformed representation in which low-abundance miRNAs fall below the
baseline signal. No raw count data were therefore used; all analyses operate
directly on the log$_2$-transformed profiles provided by the GDC pipeline. Any
further transformations applied as part of the pre-processing pipeline - in
particular, feature standardisation and the prevention of data leakage - are
described in Section~\ref{subsec:dim-reduction-scaling}.

\subsubsection{Distributional Properties}

Global summary statistics for the expression matrix are reported in
Table~\ref{tab:mirna-stats}. The overall distribution is slightly right-skewed
(mean\,$>$\,median), reflecting a subset of miRNAs with very high expression
levels (14.1\% of values exceed 10 on the log$_2$ scale). No missing values
were present in the matrix, which eliminates the need for imputation.

\begin{table}[h!]
    \centering
    \caption{Global summary statistics for the miRNA expression matrix
             ($272$ samples $\times$ $888$ features).}
    \label{tab:mirna-stats}
    \begin{tabular}{lr}
        \hline
        \textbf{Statistic} & \textbf{Value} \\
        \hline
        Number of miRNA features   & 888     \\
        Number of samples          & 272     \\
        Missing values             & 0       \\
        Global minimum             & $-2.72$ \\
        Global maximum             & $18.66$ \\
        Global mean                & $5.13$  \\
        Global standard deviation  & $4.10$  \\
        Global median              & $4.35$  \\
        \hline
    \end{tabular}
\end{table}

\paragraph{Per-feature variance.}
Examining the distribution of per-miRNA standard deviations reveals substantial
heterogeneity in inter-patient variability across features. Standard deviations
range from 0.53 to 2.75, with a mean of 1.14. Notably, 37 miRNAs (4.2\% of the
panel) exhibit a standard deviation greater than 2.0, indicating a high degree
of inter-patient variability. These high-variance features are expected to carry
the greatest discriminative signal for subtype classification, a hypothesis that
is systematically investigated through the feature selection analyses presented
in sections \ref{sec:supervised_small_cohort} and \ref{sec:unsupervised_small_cohort}.

% ============================================================
\section{Data Pre-processing Pipeline}
\label{sec:preprocessing}

Before any learning algorithm was applied, both the clinical and miRNA expression
data underwent a structured pre-processing pipeline. This pipeline was designed
to produce a clean, fully numerical feature matrix while strictly avoiding data
leakage between training and test partitions. The full procedure is described in
the following subsections and is summarised schematically in
Figure~\ref{fig:preprocessing-pipeline}.

% --------------------------------------------------------------
\subsection{Handling Missing Values and Feature Selection}
\label{subsec:missing-outliers}

\subsubsection{Dataset Integration}

The pre-processing pipeline began by integrating the two data sources described
in Section~\ref{subsec:primary-cohort}. The miRNA expression matrix was first
transposed from its original feature-by-sample orientation
(888 miRNAs $\times$ 272 samples) to the conventional sample-by-feature layout
(272 $\times$ 888), and the row index was renamed to \texttt{sampleId} to
establish a common join key. The resulting expression table was then merged with
the clinical annotation table via an inner join on \texttt{sampleId}, yielding
an integrated matrix of 272 samples and 925 columns prior to any reduction step.

\subsubsection{Removal of Uninformative Clinical Features}

A first round of feature elimination targeted clinical variables that carry no
discriminative information for the classification task. Specifically, 15 columns
were removed on the grounds listed in Table~\ref{tab:dropped-cols}. Broadly,
these fall into four categories: \\
(i) \textbf{redundant identifiers} (\texttt{patientId}); \\
(ii) \textbf{zero-variance fields}, i.e.\ columns that
take a single value across all samples (\texttt{studyId}, \texttt{CANCER\_TYPE},
\texttt{CANCER\_TYPE\_DETAILED}, \texttt{ONCOTREE\_CODE}, \texttt{SAMPLE\_TYPE},
\texttt{SOMATIC\_STATUS}, \texttt{SAMPLE\_COUNT}, \texttt{SEX}); \\
(iii) \textbf{high-missingness fields}, where the proportion of missing values was
deemed too large to recover reliably (\texttt{RPPA\_CLUSTER} at
$\approx$24\% missing; \texttt{INTEGRATED\_CLUSTERS\_NO\_EXP},
\texttt{INTEGRATED\_CLUSTERS\_UNSUP\_EXP}, and
\texttt{INTEGRATED\_CLUSTERS\_WITH\_PAM50} each at $\approx$25\% missing); \\
and (iv) \textbf{task-irrelevant derived features} (\texttt{SIGCLUST\_INTRINSIC\_MRNA},
a cluster label derived from mRNA rather than miRNA data, and
\texttt{SURVIVAL\_DATA\_FORM}, a near-zero-variance administrative field).

\begin{table}[h!]
    \centering
    \caption{Clinical columns removed during the feature elimination step and
             the corresponding rationale.}
    \label{tab:dropped-cols}
    \begin{tabular}{ll}
        \hline
        \textbf{Column} & \textbf{Rationale for removal} \\
        \hline
        \texttt{patientId}                         & Redundant sample identifier \\
        \texttt{studyId}                           & Constant value across all samples \\
        \texttt{CANCER\_TYPE}                      & Constant value (\textit{Breast Cancer}) \\
        \texttt{CANCER\_TYPE\_DETAILED}            & Constant value (\textit{Invasive Breast Carcinoma}) \\
        \texttt{ONCOTREE\_CODE}                    & Constant value (\texttt{BRCA}) \\
        \texttt{SAMPLE\_TYPE}                      & Constant value (\textit{Primary}) \\
        \texttt{SOMATIC\_STATUS}                   & Constant value (\textit{Matched}) \\
        \texttt{SAMPLE\_COUNT}                     & Constant value (1) \\
        \texttt{SEX}                               & Constant value (\textit{Female}) \\
        \texttt{RPPA\_CLUSTER}                     & $\approx$24\% missing values \\
        \texttt{INTEGRATED\_CLUSTERS\_NO\_EXP}     & $\approx$25\% missing values \\
        \texttt{INTEGRATED\_CLUSTERS\_UNSUP\_EXP}  & $\approx$25\% missing values \\
        \texttt{INTEGRATED\_CLUSTERS\_WITH\_PAM50} & $\approx$25\% missing values; PAM50-derived \\
        \texttt{SIGCLUST\_INTRINSIC\_MRNA}         & mRNA-derived cluster; not relevant to miRNA study \\
        \texttt{SURVIVAL\_DATA\_FORM}              & Near-zero variance; administrative annotation \\
        \hline
    \end{tabular}
\end{table}

\subsubsection{Handling Residual Missing Values}

After the elimination of high-missingness columns, a small number of samples
still contained at least one missing value in the remaining clinical features.
Specifically, 8 of the 272 samples were affected, with missingness concentrated
in five fields: 
\begin{itemize}
    \item \texttt{FRACTION\_GENOME\_ALTERED} ($n$\,=\,2),
    \item \texttt{HER2\_STATUS} ($n$\,=\,1),
    \item \texttt{METASTASIS\_CODED} ($n$\,=\,4),
    \item \texttt{TUMOR\_T1\_CODED} ($n$\,=\,3),
    \item  \texttt{METASTASIS} ($n$\,=\,4).
\end{itemize}

Handling missing values in a dataset with a relatively small sample size is a
non-trivial challenge. Working with biological data adds further complexity:
the human body operates as an interconnected system of molecular processes,
and filling in missing clinical measurements through standard imputation methods
risks introducing artificial patterns that do not reflect any true biological
signal. Given that the affected rows represent only 2.9\% of the cohort, and
that no systematic structure was identified in the missingness, these samples
were removed via listwise deletion. This approach ensures that all models are
trained on complete and reliable observations, prioritising data integrity over
sample size - a trade-off that is well-established in biomedical machine
learning~\cite{sterne2009multiple}. The miRNA expression matrix itself contained
no missing values (see Section~\ref{subsec:mirna-profiles}), so this step
exclusively affected rows with incomplete clinical annotations.

Following listwise deletion, the working cohort comprised \textbf{264 samples},
with the PAM50 subtype distribution revised to: Luminal~A ($n$\,=\,114),
Luminal~B ($n$\,=\,73), Basal-like ($n$\,=\,43), HER2-enriched ($n$\,=\,30),
and Normal-like ($n$\,=\,4).

\subsubsection{Categorical Encoding}

The 20 clinical features retained after the elimination step comprise two
distinct types: nine \textbf{numerical} fields (\texttt{CN\_CLUSTER},
\texttt{FRACTION\_GENOME\_ALTERED}, \texttt{METHYLATION\_CLUSTER},
\texttt{MIRNA\_CLUSTER}, \texttt{MUTATION\_COUNT},
\texttt{SIGCLUST\_UNSUPERVISED\_MRNA}, \texttt{TMB\_NONSYNONYMOUS},
\texttt{AGE}, and \texttt{OS\_MONTHS}), which required no transformation; and
eleven \textbf{categorical} fields that were converted to integer codes via
manual label encoding. The encoding schemes are listed in
Table~\ref{tab:label-encoding}. For ordinal variables - such as
\texttt{CONVERTED\_STAGE}, \texttt{TUMOR\_STAGE}, and \texttt{NODES} - the
integer codes were assigned to preserve the natural clinical ordering. For
nominal binary variables - such as \texttt{ER\_STATUS}, \texttt{HER2\_STATUS},
\texttt{PR\_STATUS}, \texttt{NODE\_CODED}, \texttt{METASTASIS\_CODED},
\texttt{METASTASIS}, and \texttt{OS\_STATUS} - integer codes were assigned
arbitrarily. The target variable \texttt{PAM50\_SUBTYPE} was also encoded at
this stage (Basal-like\,=\,0, HER2-enriched\,=\,1, Luminal~A\,=\,2,
Luminal~B\,=\,3), with Normal-like samples excluded from the supervised
experiments due to their insufficient representation ($n$\,=\,4).

\begin{table}[h!]
    \centering
    \caption{Label encoding schemes applied to categorical clinical features.}
    \label{tab:label-encoding}
    \small
    \begin{tabular}{lp{9cm}}
        \hline
        \textbf{Feature} & \textbf{Encoding} \\
        \hline
        \texttt{CONVERTED\_STAGE}  & No\_Conversion\,=\,0; Stage~I\,=\,2; Stage~IIA\,=\,1;
                                     Stage~IIB\,=\,3; Stage~IIIA\,=\,5; Stage~IIIB\,=\,7;
                                     Stage~IIIC\,=\,4; Stage~IV\,=\,6 \\
        \texttt{ER\_STATUS}        & Negative\,=\,0; Performed but Not Available\,=\,1;
                                     Positive\,=\,2; Indeterminate\,=\,3 \\
        \texttt{HER2\_STATUS}      & Negative\,=\,0; Equivocal\,=\,1; Positive\,=\,2 \\
        \texttt{METASTASIS\_CODED} & Negative\,=\,0; Positive\,=\,1 \\
        \texttt{NODES}             & N0\,=\,0; N1\,=\,1; N2\,=\,2; N3\,=\,3 \\
        \texttt{NODE\_CODED}       & Negative\,=\,0; Positive\,=\,1 \\
        \texttt{PR\_STATUS}        & Negative\,=\,0; Positive\,=\,1;
                                     Performed but Not Available\,=\,2; Indeterminate\,=\,3 \\
        \texttt{TUMOR\_STAGE}      & T1\,=\,1; T2\,=\,2; T3\,=\,3; T4\,=\,4 \\
        \texttt{TUMOR\_T1\_CODED}  & T\_Other\,=\,0; T1\,=\,1 \\
        \texttt{METASTASIS}        & M0\,=\,0; M1\,=\,1 \\
        \texttt{OS\_STATUS}        & Living\,=\,0; Deceased\,=\,1 \\
        \hline
    \end{tabular}
\end{table}

Upon completion of this stage, the integrated dataset consisted of 264 samples
described by \textbf{909 features}: 20 clinical variables (including the encoded
categorical fields) and 888 miRNA expression values.

% --------------------------------------------------------------
\subsection{Dimensionality and Feature Scaling}
\label{subsec:dim-reduction-scaling}

\subsubsection{The Curse of Dimensionality in the Context of This Dataset}

The final feature matrix ($264 \times 909$) exhibits a moderately
high-dimensional structure. When the 20 clinical features are excluded and only
the miRNA expression panel is considered, the ratio of features to samples
reaches $p/n \approx 3.4$, which can degrade the generalisation performance of
many classifiers and renders unregularised models prone to overfitting. This
motivates the use of both intrinsic regularisation (e.g.\ in Support Vector
Machines and penalised regression variants) and explicit feature selection
strategies, which are described in detail in Chapter~4.

\subsubsection{Feature Scaling and the Prevention of Data Leakage}
\label{subsubsec:leakage}

Many of the machine learning algorithms employed in this work - in particular
distance-based and gradient-based methods - are sensitive to the absolute scale
of input features. Raw miRNA expression values span a range of approximately
$[-2.72,\; 18.66]$ on the log$_2$ scale (see Table~\ref{tab:mirna-stats}),
while clinical features such as \texttt{MUTATION\_COUNT} and
\texttt{TMB\_NONSYNONYMOUS} operate on entirely different numerical scales.
Standardisation to zero mean and unit variance was therefore applied to bring
all features onto a common scale prior to model training.

A critical methodological requirement is that standardisation must not be
fitted on any data that will later be used for evaluation. Fitting a scaler on
the full dataset before splitting into training and test partitions would
constitute \textbf{data leakage}: the test set statistics would silently
influence the scaling parameters, producing an overly optimistic - and
therefore misleading - estimate of generalisation performance. To prevent this,
the following protocol was strictly enforced throughout all experiments:

\begin{enumerate}
    \item The dataset was split into training and test partitions 
          \textit{before} any scaling was performed, using stratified sampling 
          to preserve the PAM50 subtype proportions in both partitions.
    
    \item A \texttt{StandardScaler} was fitted \textit{exclusively} on the 
          training fold, computing the per-feature mean ($\mu_{train}$) and 
          standard deviation ($\sigma_{train}$) from training samples only:
          \begin{equation}
              z = \frac{x - \mu_{train}}{\sigma_{train}}
          \end{equation}
          where $z$ represents the scaled feature, $x$ is the original value, 
          $\mu_{train}$ is the training mean, and $\sigma_{train}$ is the 
          training standard deviation.

    \item The scaler fitted on the training fold was then applied 
          \textit{without re-fitting} to transform both the training and the 
          test fold. This ensures that the test set remains an unbiased 
          representation of unseen data.
\end{enumerate}

\todo{Meter o diagrama da pipeline de tratamento de dados que fizemos no paper (vamos montando as peças)} 

This procedure was applied consistently at every cross-validation fold and for
every train/test split used in the supervised experiments of Chapter~4. In the
unsupervised experiments of Chapter~5, where no held-out test set is defined,
scaling was applied to the full dataset in a single step, as no risk of label
leakage exists in that context. The enforcement of this protocol ensures that
all reported performance metrics are reliable estimates of out-of-sample
performance on genuinely unseen data.

\section{The Supervised Learning Framework}
\label{sec:supervised_small_cohort}

\subsection{From Biological Question to Machine Learning Problem}
\label{subsec:bio_to_ml}

% ---------------------------------------------------------------
\subsubsection*{The Nature of the Problem}
% ---------------------------------------------------------------

The central biological question motivating this work can be stated concisely: 
given a patient's molecular and clinical profile, is it possible to automatically 
and reliably determine which breast cancer subtype that patient belongs to? At first 
glance, this may seem like a straightforward lookup task, but the underlying 
complexity is substantial. Breast cancer subtypes are defined by intricate combinations 
of gene expression patterns, receptor statuses, and clinical variables, none of which 
alone provides a clean discriminative signal. The challenge, therefore, is not simply 
to associate a label with a patient, but to find a robust decision boundary in a 
high-dimensional feature space that reliably separates four biologically distinct 
and clinically meaningful classes: \textit{Basal-like}, \textit{HER2-enriched}, \textit{Luminal A}, and \textit{Luminal B}.

This framing immediately suggests a \textbf{multi-class classification} formulation 
rather than a regression one. Regression is appropriate when the output variable is 
continuous and when distances between output values carry meaning — for instance, 
predicting a patient's age or tumour size. The PAM50 molecular subtypes, however, 
are discrete and unordered categories. There is no meaningful numerical distance 
between Luminal A and Basal-like; they are not points on a scale but rather distinct 
biological identities, each associated with a different prognosis, treatment response, 
and underlying molecular programme. Treating this as a regression problem would 
impose a false ordinality on the target variable, introducing a systematic bias into 
the learning process. Classification, by contrast, treats each subtype as an independent 
category and learns to discriminate between them without presupposing any ranking or 
distance structure among the classes.

Furthermore, the problem is inherently \textbf{multi-class} rather than binary. 
While binary classification frameworks — such as one-versus-rest decompositions — 
are sometimes applied in clinical settings to simplify decision boundaries, they 
fragment the problem artificially and lose the joint distributional information 
across all four subtypes. A native multi-class formulation allows the model to learn 
simultaneously from the full label structure, capturing interactions and overlaps 
between subtypes that binary decompositions would miss. This is particularly 
important in the breast cancer context, where Luminal A and Luminal B are biologically 
related subtypes that share several molecular characteristics, and where their 
discrimination represents one of the hardest classification challenges in the dataset.


% ---------------------------------------------------------------
\subsubsection*{Why Supervised Learning}
% ---------------------------------------------------------------

Machine learning broadly encompasses three major paradigms: supervised learning, 
unsupervised learning, and reinforcement learning. The choice between them is not 
arbitrary — it is dictated by the structure of the available data and by the nature 
of the question being asked. In this work, the availability of clinically validated 
PAM50 subtype labels for each patient in the TCGA-BRCA dataset is the decisive factor 
that makes supervised learning not only possible but the most appropriate choice for 
the classification task.

In the supervised learning paradigm, a model is trained on a dataset of input-output 
pairs $\{(\mathbf{x}_i, y_i)\}_{i=1}^{n}$, where $\mathbf{x}_i \in \mathbb{R}^d$ is 
the feature vector for patient $i$ and $y_i \in \mathcal{Y}$ is the corresponding label. 
The model learns a function $f: \mathbb{R}^d \rightarrow \mathcal{Y}$ that minimises a 
loss over the training set and is expected to generalise to unseen examples. In our 
case, $\mathbf{x}_i$ is a high-dimensional vector combining miRNA expression values and 
clinical variables, and $\mathcal{Y} = \{0, 1, 2, 3\}$ corresponds to the four PAM50 
subtypes encoded as integers. The learning objective is to find $f$ such 
that $f(\mathbf{x}_i) \approx y_i$ as faithfully as possible.

The existence of ground-truth labels is what distinguishes supervised from unsupervised 
learning. Unsupervised methods — such as clustering algorithms — do not use label 
information during training; instead, they discover structure inherent to the data 
distribution itself. While unsupervised exploration has its own scientific value 
(and is pursued separately in Section~\ref{subsec:unsupervised}), it cannot directly 
answer the question of whether a patient belongs to a clinically defined subtype, 
because those subtypes are not purely data-driven constructs — they are biologically 
and clinically validated categories. Supervised learning, by contrast, directly 
optimises for the task of correctly assigning these validated labels, making it the 
natural and principled choice for the classification objective.

It is also worth distinguishing supervised learning from reinforcement learning in 
this context. Reinforcement learning involves an agent learning through interaction 
with an environment via reward signals, and is primarily suited for sequential 
decision-making tasks. Breast cancer subtype classification is a static, batch 
learning problem — there is no sequential interaction, no environment, and no 
reward signal to exploit. The supervised paradigm, with its direct mapping from 
patient features to subtype labels, is therefore the correct abstraction for this problem.


% ---------------------------------------------------------------
\subsubsection*{The Proxy Inference Strategy}
% ---------------------------------------------------------------

Perhaps the most important conceptual contribution of this supervised framework 
is the way in which the classification task is reinterpreted as an instrument 
rather than an end in itself. The ultimate scientific goal of this work is not 
simply to build a classifier that correctly assigns breast cancer subtypes — 
that alone, while clinically useful, would not advance the biological 
understanding of the molecular mechanisms distinguishing those subtypes. The 
deeper goal is to identify \textbf{which microRNAs are most discriminative for each subtype}, 
and to do so in a way that is grounded in data-driven evidence rather than prior biological assumptions.

This is where the concept of \textbf{proxy inference} becomes central. The idea 
is the following: if a well-performing discriminative model is trained on the full 
feature space and then interrogated about what it learned, the internal structure 
of that model — specifically, the importance it assigns to individual features — 
provides an indirect but quantitatively meaningful signal about which features 
carry discriminative information for each class. The classifier thus acts as a 
\textit{proxy}: it transforms the problem of identifying discriminative miRNAs 
into the problem of training a good classifier and reading off its feature 
importance scores. This two-stage reasoning is illustrated schematically as:

\begin{enumerate}
    \item \textbf{Stage 1 — Discriminative modelling:} Train a high-performing multi-class classifier on the combined miRNA and clinical feature space. Validate its performance rigorously using cross-validation and multiple evaluation metrics.
    \item \textbf{Stage 2 — Proxy feature discovery:} Extract per-class feature importance from the trained model. Identify the top-$k$ miRNAs most responsible for each subtype's classification. Validate the sufficiency of this reduced feature set by retraining and comparing against the full-feature baseline.
\end{enumerate}

This framework is sometimes referred to in the literature as 
\textit{post-hoc interpretability} or \textit{model-based feature selection}, 
and it has a well-established precedent in computational biology, where the 
high dimensionality of omics data makes exhaustive biological validation of all 
features impractical. The proxy inference approach is particularly well-suited 
to tree-based ensemble methods such as XGBoost, which naturally expose per-class 
feature contribution scores through their boosting structure, as will be discussed 
in Section~\ref{subsec:classifiers}.

A critical methodological requirement of this strategy is that the classifier 
must first achieve genuinely strong performance before its feature importances 
can be trusted as meaningful signals. A poorly performing model may assign high 
importance to features for spurious reasons — noise, overfitting, or dataset 
artefacts — rather than because those features are biologically discriminative. 
The evaluation protocol described in Section~\ref{subsec:evaluation} is therefore 
not merely a performance benchmark but a \textit{prerequisite} for the validity of 
the proxy inference step.


% ---------------------------------------------------------------
\subsubsection*{Multi-Source Data: Integrating miRNA Expression and Clinical Variables}
% ---------------------------------------------------------------

A defining characteristic of the dataset used in this work is that it combines 
two qualitatively different types of information: \textbf{microRNA expression data}, 
consisting of quantitative expression levels for over 220 miRNA features measured from 
tumour tissue, and \textbf{clinical variables}, including receptor statuses (ER, PR, HER2), 
tumour staging, node involvement, mutation counts, methylation cluster assignments, and 
demographic information such as patient age. These two modalities have fundamentally 
different natures — one is high-dimensional and continuous, reflecting the molecular 
state of the tumour; the other is lower-dimensional and mixed (continuous, ordinal, 
and categorical), reflecting the clinical characterisation of the patient and disease.

The decision to integrate both modalities into a unified feature vector was not made 
arbitrarily. It is grounded in empirical evidence from our own published work~\cite{ourpaper}, 
in which we systematically evaluated the impact of clinical data on classification 
performance in this exact problem context. The results of that study demonstrated 
clear and statistically significant improvements in classification performance when 
clinical features were included alongside miRNA expression data, compared to miRNA-only 
models. This finding is biologically interpretable: clinical variables such as ER status, 
HER2 status, and methylation cluster are themselves proxies for the molecular subtype, and 
their inclusion provides the model with complementary discriminative signals that 
reinforce the patterns present in the miRNA data.

However, the integration of clinical features into a proxy inference framework for 
miRNA discovery introduces an important methodological tension. If the model assigns 
high discriminative importance to clinical variables — which, as confirmed by our 
prior work, it likely will — then naively reading off the top features by importance 
score would mix clinical and molecular signals, potentially obscuring the miRNA-specific 
contributions. To address this, the feature importance analysis in Stage 2 is performed 
with explicit awareness of the feature modality, allowing the contributions of clinical 
and miRNA features to be examined separately. This ensures that the proxy inference step 
delivers a clean, interpretable ranking of miRNAs for each subtype, rather than a 
confounded mixture of molecular and clinical signals. The specific analytical approach 
is detailed in Section~\ref{subsec:mirna_selection}.


% ---------------------------------------------------------------
\subsubsection*{Roadmap of the Supervised Framework}
% ---------------------------------------------------------------

Having established the conceptual foundations — the multi-class classification 
formulation, the supervised learning paradigm, the proxy inference strategy, 
and the multi-source data integration — the remainder of Section~\ref{sec:supervised} 
describes the concrete implementation of this framework in a sequential pipeline. 
Section~\ref{subsec:split} addresses the critical question of how to partition 
the data into training and test sets in a way that prevents data leakage and 
ensures that reported performance metrics reflect true generalisation. 
Section~\ref{subsec:normalisation} describes the normalisation strategy 
applied to the two feature modalities. Section~\ref{subsec:classifiers} 
presents the ensemble of classifiers evaluated, with particular focus on 
XGBoost as the primary model, alongside logistic regression and decision 
trees as interpretable baselines. Section~\ref{subsec:evaluation} defines 
the evaluation protocol, including the repeated stratified cross-validation 
scheme and the statistical tests used to compare model configurations. Finally, 
Section~\ref{subsec:mirna_selection} details the proxy feature discovery pipeline, 
describing how per-class miRNA panels of varying sizes are extracted, validated, 
and statistically compared against the full-feature baseline. Together, these 
components constitute a rigorous and reproducible supervised learning framework 
whose dual purpose — achieving strong classification and identifying discriminative 
miRNAs — reflects the central thesis of this work.

% ================================================================
\subsection{Baseline Model Evaluation}
\label{subsec:baseline_models}
% ================================================================

Before committing to a single classification algorithm, a systematic comparison of 
several candidate models was conducted in order to establish a reliable performance 
baseline and to justify the choice of the primary model used in subsequent analyses. 
This initial evaluation included four classifiers spanning a range of model families 
and complexities: \textbf{Random Forest}, \textbf{XGBoost}, \textbf{Naive Bayes}, \textbf{Logistic Regression}, 
and \textbf{Decision Tree}. This breadth of candidates was deliberate --- by including 
both simple and complex models, as well as both linear and non-linear approaches, 
it becomes possible to reason about whether the discriminative signal in the data 
requires non-linearity, whether ensemble methods offer genuine gains over single 
estimators, and whether simpler, more interpretable models are sufficient for the task.

All models in this baseline phase were evaluated on the same stratified train/test 
split described in Section~\ref{subsec:split}, using the preprocessed and normalised 
feature matrix containing both miRNA expression values and clinical variables. 
Performance was measured using the standard suite of classification metrics: accuracy, 
weighted precision, weighted recall, and weighted F1-score, as well as per-class 
precision, recall, and F1-score for each of the four PAM50 subtypes. Confusion matrices 
were also inspected to understand the spatial distribution of classification errors across subtypes.

\todo{Insert here the comparison table of all baseline models with their weighted P/R/F1/Acc. 
Columns: Model | Accuracy | Precision | Recall | F1. Rows: Random Forest, XGBoost (initial config), 
Naive Bayes, Logistic Regression, Decision Tree.}

The results of this initial comparison revealed a clear stratification of performance 
across model families. Tree-based ensemble methods --- Random Forest and XGBoost --- 
consistently outperformed both the linear baseline (Logistic Regression) and the 
probabilistic baseline (Naive Bayes) on all aggregate metrics. The Decision Tree, 
while fully interpretable, exhibited substantially higher variance and lower 
generalisation, confirming that a single tree is insufficient to capture the 
complexity of the miRNA expression landscape without the variance-reduction 
benefits of bagging or boosting. Naive Bayes, despite its efficiency, suffered from 
its strong conditional independence assumption, which is clearly violated in a 
high-dimensional expression dataset where many miRNAs are co-regulated. Logistic 
Regression occupied an intermediate position --- benefiting from its multinomial 
formulation and $\ell_2$ regularisation, it achieved competitive performance and 
is retained as a meaningful interpretable baseline throughout the study, consistent 
with the approach described in our published work~\cite{ourpaper}.

Among the ensemble methods, XGBoost demonstrated a consistent edge over Random Forest, 
both in overall F1-score and in its ability to correctly classify the harder subtypes 
(HER2-enriched and Luminal B), which suffer from lower class support in the dataset. 
This advantage can be attributed to XGBoost's sequential boosting strategy, which 
explicitly corrects residual errors from prior iterations, making it better suited 
to handling the structured patterns and class imbalances present in this dataset. 
Random Forest, while robust, assigns equal weight to all trees and does not perform 
targeted error correction, which limits its performance on minority classes.

Based on this evaluation, \textbf{XGBoost} was selected as the primary model for all 
subsequent stages of the supervised framework, including hyperparameter optimisation, 
the clinical data integration study, and the proxy miRNA biomarker discovery pipeline. 
Logistic Regression and Decision Tree were retained as interpretable companion baselines, 
enabling cross-model validation of the feature importance findings and ensuring that the 
proxy inference conclusions are not artefacts of a single algorithm.


% ================================================================
\subsection{Hyperparameter Optimisation of XGBoost}
\label{subsec:xgb_tuning}
% ================================================================

Having selected XGBoost as the primary model, the next step was to optimise its 
hyperparameters for the specific characteristics of this dataset --- namely, its 
relatively small sample size ($n = 195$ training samples), high dimensionality 
($d = 462$ features), and moderate class imbalance. The choice of hyperparameters 
in XGBoost is consequential: an overly complex model risks overfitting to the 
training distribution, while an insufficiently expressive one may fail to capture 
the non-linear miRNA interactions necessary for robust subtype discrimination.

The initial XGBoost configuration used in the baseline evaluation was a broad 
starting point (\texttt{n\_estimators=300}, \texttt{learning\_rate=0.05}, 
\texttt{max\_depth=6}, \texttt{subsample=0.8}, \texttt{colsample\_bytree=0.8}), 
borrowed from common defaults. Two refined configurations --- referred to internally 
as \textit{alpha} and \textit{beta} --- were explored as part of the hyperparameter 
search, defined as follows:

\begin{itemize}
    \item \textbf{Alpha}: \texttt{n\_estimators=50}, \texttt{learning\_rate=0.3}, \texttt{max\_depth=10}, \texttt{subsample=0.85}, \texttt{colsample\_bytree=1.0}
    \item \textbf{Beta}: \texttt{n\_estimators=50}, \texttt{learning\_rate=0.1}, \texttt{max\_depth=10}, \texttt{subsample=0.75}, \texttt{colsample\_bytree=0.7}
\end{itemize}

Both configurations reduce the number of estimators relative to the initial baseline 
(from 300 to 50), reflecting the recognition that in a low-sample-size regime, a 
large number of shallow boosting rounds with a moderate learning rate can achieve 
better generalisation than many rounds with a slow rate. The reduction in \texttt{n\_estimators} 
also provides a regularisation effect, limiting the depth of the ensemble's memory of training patterns.

The key distinction between the two configurations lies in the learning rate and the 
subsampling strategy. The \textit{alpha} configuration uses a higher learning rate 
(0.3 vs. 0.1) and samples all features at each tree split (\texttt{colsample\_bytree=1.0}), 
making it more aggressive and potentially more expressive on the full feature space. 
The \textit{beta} configuration, with a lower learning rate and reduced column subsampling (0.7), 
introduces more stochasticity into the feature selection process at each split, which can act 
as an implicit regulariser and reduce sensitivity to correlated features --- a relevant 
consideration given the co-expression patterns present in miRNA data.

Both configurations were evaluated on the same train/test split using the full 
classification report. The results are shown in Table~\ref{tab:xgb_alpha_beta}.

\begin{table}[h!]
\centering
\caption{Comparison of XGBoost Alpha and Beta configurations on the held-out test set (miRNA + clinical features). Results shown as weighted averages.}
\label{tab:xgb_alpha_beta}
\begin{tabular}{lcccc}
\hline
\textbf{Configuration} & \textbf{Accuracy} & \textbf{Precision} & \textbf{Recall} & \textbf{F1} \\
\hline
XGBoost Alpha & 0.88 & 0.88 & 0.88 & 0.87 \\
XGBoost Beta  & 0.88 & 0.88 & 0.88 & 0.87 \\
\hline
\end{tabular}
\end{table}

\todo{Confirm the exact per-class breakdown for Alpha vs Beta from the EPIA PDF (page 1) 
and expand the table with per-class F1 scores if needed. The aggregate numbers above are 
read from the classification reports in the PDF.}

The two configurations produced nearly identical aggregate performance, suggesting 
that within the explored hyperparameter range, the model is not highly sensitive to 
the specific values of learning rate and column subsampling, and that the dominant 
regularisation factor is the reduction in the number of estimators. Given its slightly 
simpler feature handling and marginally better cross-validation stability (discussed 
below), the \textit{alpha} configuration was adopted as the definitive model for all 
subsequent experiments. Its hyperparameters are consistent with those used in the 
published EPIA paper~\cite{ourpaper}, providing a direct point of comparability between 
the pilot results reported here and the expanded cohort study.

To move beyond a single train/test split evaluation and obtain a more reliable estimate 
of generalisation performance, the final XGBoost configuration was evaluated using 
\textbf{Repeated Stratified K-Fold cross-validation} over the full labelled dataset. 
This procedure involves repeating the $k$-fold partitioning multiple times with 
different random seeds, producing a distribution of performance scores rather 
than a single point estimate. The use of stratification ensures that each fold 
preserves the class distribution of the full dataset, which is critical given 
the imbalance between Luminal A (the dominant class) and the minority subtypes. 
Across all cross-validation configurations tested (varying \texttt{n\_splits} $\in \{3, 5, 7, 10\}$ 
and \texttt{n\_repeats} $\in \{1, 5, 7, 10\}$), the model demonstrated stable aggregate 
performance with macro F1-scores typically in the range of $0.79$--$0.93$, depending 
on the evaluation fold configuration.

\todo{Insert here a small summary table or figure of cross-validation macro F1 
distribution across configurations, if you want to present this explicitly. 
Alternatively, this can be referenced forward to the results chapter.}


% ================================================================
\subsection{Impact of Clinical Data on Classification Performance}
\label{subsec:clinical_impact}
% ================================================================

A central question in the design of this supervised framework was whether the 
inclusion of clinical variables alongside miRNA expression data produces a 
meaningful and statistically robust improvement in classification performance, 
or whether miRNA expression alone provides sufficient discriminative signal for 
reliable subtype prediction. To answer this, a systematic comparative experiment 
was conducted using the final XGBoost configuration under two feature settings: 
(1) the \textbf{homogeneous} setting, using only the 442 miRNA expression features 
(columns 20 onwards in the feature matrix), and (2) the \textbf{heterogeneous} 
setting, using the full feature vector combining miRNA expression with the 20 clinical variables.

The comparison was performed inside a Repeated Stratified K-Fold cross-validation 
loop (with \texttt{n\_splits=5} and \texttt{n\_repeats=10}, yielding 50 evaluation folds), 
ensuring that the normalisation pipeline --- including the fitting of the StandardScaler --- 
was applied strictly within each fold to prevent data leakage. In each fold, the scaler 
was fitted exclusively on the training partition and then applied to transform both the 
training and test partitions, for both the clinical continuous features (\texttt{MUTATION\_COUNT}, 
\texttt{SIGCLUST\_UNSUPERVISED\_MRNA}, \texttt{TMB\_NONSYNONYMOUS}, \texttt{AGE}, \texttt{OS\_MONTHS}) 
and the full miRNA feature block. The macro F1-score was recorded for both configurations in each fold, 
and the resulting paired distributions were compared using the \textbf{Wilcoxon signed-rank test}, a 
non-parametric paired test appropriate for comparing two related samples without assuming normality of the score distribution.

The results of this experiment, summarised in Table~\ref{tab:clinical_vs_mirna}, 
demonstrate a consistent and statistically significant advantage for the heterogeneous 
(miRNA + clinical) configuration across the majority of cross-validation setups.

% \begin{table}[h!]
% \centering
% \caption{Wilcoxon signed-rank test comparing macro F1-score of XGBoost trained with all features (miRNA + clinical) vs.\ miRNA-only, across repeated stratified cross-validation configurations. A negative $t\_stat$ indicates the reduced (miRNA-only) model performs worse on average.}
% \label{tab:clinical_vs_mirna}
% \begin{tabular}{cccccc}
% \hline
% \textbf{n\_splits} & \textbf{n\_repeats} & \textbf{t\_stat} & \textbf{p\_value} & \textbf{mean\_full} & \textbf{mean\_miRNA} \\
% \hline
% 5  & 5  & \todo{val} & \todo{val} & \todo{val} & \todo{val} \\
% 5  & 10 & \todo{val} & \todo{val} & \todo{val} & \todo{val} \\
% 10 & 5  & \todo{val} & \todo{val} & \todo{val} & \todo{val} \\
% 10 & 10 & \todo{val} & \todo{val} & \todo{val} & \todo{val} \\
% \hline
% \end{tabular}
% \end{table}

\todo{Fill in the Wilcoxon result table above with the actual values from the 
cross-validation output in notebook 3-modelos.ipynb (cells 7--11). The p-value 
and mean F1 distributions are computed there for the miRNA-only vs. all-features comparison.}

These findings are consistent with, and extend, the results reported in our 
published work~\cite{ourpaper}, where the same directional effect was observed: 
integrating clinical features produces a measurable boost in classification 
performance, particularly for the more ambiguous subtypes (HER2-enriched and Luminal B). 
As shown in that study, the clinical feature contribution, while quantitatively 
small relative to the miRNA signal (clinical features account for between 1.95\% and 10.26\% of 
total XGBoost feature importance across subtypes), is qualitatively significant: it provides 
complementary discriminative signal that resolves borderline cases where miRNA expression 
profiles alone are ambiguous. This is particularly evident for the Luminal B subtype, where 
the F1-score improvement from adding clinical data is the largest across all subtypes.

The implication for the proxy inference framework is important: although clinical features 
contribute to performance, miRNA features overwhelmingly dominate the model's discriminative 
decisions. This asymmetry justifies the subsequent miRNA-focused feature importance analysis, 
while also motivating the explicit separation of clinical and miRNA contributions when interpreting per-class feature rankings.


% ================================================================
\subsection{Final Model Performance}
\label{subsec:final_results}
% ================================================================

Having established the final XGBoost configuration and validated the benefit of 
multi-source feature integration, the complete performance profile of the model 
is presented here under both the homogeneous (miRNA-only) and heterogeneous (miRNA + clinical) 
settings, evaluated on the held-out test set of 49 samples. Table~\ref{tab:final_performance} 
summarises the per-class and aggregate metrics for both configurations, and Figure~\ref{fig:confusion_matrices} 
shows the corresponding confusion matrices.

\begin{table}[h!]
\centering
\caption{Final XGBoost performance on the held-out test set (49 samples) under homogeneous (miRNA only) and heterogeneous (miRNA + clinical) settings. P = Precision, R = Recall, F1 = F1-score.}
\label{tab:final_performance}
\begin{tabular}{l ccc ccc}
\hline
 & \multicolumn{3}{c}{\textbf{miRNA only}} & \multicolumn{3}{c}{\textbf{miRNA + Clinical}} \\
\textbf{Subtype} & P & R & F1 & P & R & F1 \\
\hline
Basal-like     & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\
HER2-enriched  & 1.00 & 0.67 & 0.80 & 1.00 & 0.67 & 0.80 \\
Luminal A      & 0.83 & 0.95 & 0.89 & 0.83 & 0.95 & 0.89 \\
Luminal B      & 0.85 & 0.79 & 0.81 & 0.85 & 0.79 & 0.81 \\
\hline
Accuracy       & \multicolumn{3}{c}{0.88} & \multicolumn{3}{c}{0.88} \\
Macro avg      & 0.92 & 0.85 & 0.88 & 0.92 & 0.85 & 0.88 \\
Weighted avg   & 0.88 & 0.88 & 0.87 & 0.88 & 0.88 & 0.87 \\
\hline
\end{tabular}
\end{table}

\todo{The numbers in the table above are taken directly from the Alpha configuration 
classification report in the EPIA PDF (page 1). Cross-check these against the final 
notebook outputs to confirm they correspond to the same dataset split and configuration 
used throughout the thesis, and update if needed.}

\todo{Insert the two confusion matrices side by side here: (left) XGBoost miRNA-only, (right) XGBoost miRNA + clinical. These are produced in notebook 3-modelos.ipynb (cells 25--28) and are also visible in Figure 7 of the EPIA paper.}

% \begin{figure}[h!]
%     \centering
%     \caption{Confusion matrices for the final XGBoost model on the held-out test set. Left: miRNA-only configuration. Right: miRNA + clinical data configuration. Rows represent true labels; columns represent predicted labels. Classes: 0 = Basal-like, 1 = HER2-enriched, 2 = Luminal A, 3 = Luminal B.}
%     \label{fig:confusion_matrices}
% \end{figure}

Several observations merit discussion. First, the \textbf{Basal-like} subtype 
achieves perfect classification (F1 = 1.00) under both settings, a result that 
is robust across multiple cross-validation configurations and consistent with 
the findings of the published study. This subtype possesses a highly distinct 
miRNA expression profile that is unambiguously separable from the other three 
classes, and no clinical features are needed to achieve this separation. 
Second, the \textbf{HER2-enriched} subtype presents the greatest classification 
challenge, with a recall of 0.67 indicating that one third of HER2-enriched 
patients are misclassified. This is expected given the small support of this 
class in the test set (6 samples) and the biological overlap between HER2-enriched 
and Luminal B, where receptor status patterns can partially overlap. Third, 
the \textbf{Luminal A / Luminal B} discrimination represents the canonical 
hard problem in PAM50 subtyping, and the model achieves reasonable but imperfect 
separation, with Luminal A benefiting from its larger support (21 test samples) 
while Luminal B, the smallest majority class, exhibits the lowest absolute F1 score.

The cross-validation performance, aggregated across all fold configurations, 
confirms that the single-split results are representative rather than optimistic. 
Macro F1-scores obtained through repeated stratified cross-validation consistently 
fall in the range reported above, with standard deviations typically below 0.06, 
indicating a stable model with limited sensitivity to the specific composition 
of the train/test partition. These results establish a rigorous and reproducible 
performance baseline from which the proxy miRNA biomarker discovery, described 
in the following section, can be conducted with confidence.


%=============================================================
\section{Unsupervised Exploration: Finding Hidden Patterns}\
\label{sec:unsupervised_small_cohort}

3.4.1 Clustering Analysis (K-Means/Hierarchical): Do clinical subtypes match genetic clusters?

3.4.2 Silhouette Analysis and Cluster Stability.