%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% - Principais contributos da tese
% - Limitações do trabalho
% - Trabalho futuro
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\typeout{NT FILE chapter6.tex}%
\chapter{Conclusions and Future Work}
\label{cha:conclusions_and_future_work}

This research demonstrated that classical machine learning, when applied within
a cross-paradigm validation framework, can effectively characterise breast cancer
subtypes from microRNA expression data. The study identified a consistent
separability gradient across two independent cohorts: 
\mbox{Basal-like $>$ Luminal A $>$ HER2-Enriched $\approx$ Luminal B}, and
established that this ordering is a robust, data-intrinsic property rather than
a modelling artefact.

Supervised and unsupervised methods converged on the same conclusion: the
Basal-like subtype is reliably discriminated by miRNA expression alone, while
HER2-Enriched and Luminal B lack sufficient transcriptomic separability for
reliable classification without additional data modalities. Beyond classification,
the unsupervised analysis revealed internal heterogeneity within Luminal A that
the supervised approach alone could not have uncovered, underscoring the value
of combining complementary analytical paradigms.

The primary output of this work is a cross-cohort validated biomarker panel for
the Basal-like subtype - \textbf{MIR-577, MIR-934, MIR-18a, and MIR-135b} -
supported by multiple independent lines of computational evidence and positioned
for biological validation at NOVA Medical School~\cite{_Mendes2022Nanodelivery}.
For the remaining subtypes, the supplementary use of clinical features was
shown to improve classification, though the dependence on specific institutional
annotations limits the generalisability of those panels.

% ================================================================
\section{Future Work}
\label{sec:future_work}
% ================================================================

While this work established a robust computational baseline, several avenues
remain to deepen the biological insights and predictive accuracy of breast
cancer subtyping.

% ---------------------------------------------------------------
\subsection{Multi-Omic Integration}
\label{subsec:future_multiomics}
% ---------------------------------------------------------------

A primary limitation of the present study is the difficulty in separating
Luminal B and HER2-Enriched subtypes using miRNA expression alone. Future
research should transition towards a \textbf{multi-omic approach} that
combines miRNA with complementary data modalities. Integrating epigenetic
layers such as DNA methylation or copy number variation (CNV) could provide
the orthogonal signal necessary to resolve the boundaries where miRNA
signatures overlap. Beyond genomics, the incorporation of protein expression
data would allow for the direct validation of miRNA-target interactions,
bridging the gap between transcriptomic signatures and the functional
biological mechanics that drive subtype-specific tumour behaviour.

% ---------------------------------------------------------------
\subsection{Advanced Algorithmic Architectures}
\label{subsec:future_dl}
% ---------------------------------------------------------------

As biological datasets continue to scale and encompass more diverse data
modalities, the limitations of classical ML may be effectively addressed by
more complex models. Deep learning architectures (such as Autoencoders for
non-linear dimensionality reduction or Graph Neural Networks for modelling
molecular interaction networks) could capture dependencies that linear or
tree-based classifiers miss. Simultaneously, evolutionary and genetic algorithms
for feature selection could further refine the search for minimal clinical
panels, identifying the most potent biomarker combinations across
high-dimensional multi-omic spaces without relying on predefined assumptions
about feature relevance.

% ---------------------------------------------------------------
\subsection{Experimental Validation}
\label{subsec:future_validation}
% ---------------------------------------------------------------

The computational findings of this dissertation do not exist in isolation
from experimental biology. The cross-cohort validated Basal-like panel
(MIR-577, MIR-934, MIR-18a, and MIR-135b) is currently positioned for
\textit{in vitro} and \textit{in vivo} testing at NOVA Medical School, in
collaboration with the research team of Dr.\ Bárbara Mendes~\cite{_Mendes2022Nanodelivery}.
Future studies should prioritise these functional assays to confirm the
biological relevance of the identified miRNAs and to establish the minimum
effective panel size for a reliable diagnostic readout. If validated
experimentally, these candidates could form the basis of a targeted
diagnostic tool for Basal-like (triple-negative) breast cancer, a subtype
for which robust molecular biomarkers are particularly needed, given the
absence of hormone receptor targets that would otherwise guide therapy.
